### 2016-04-14 optimization.tex

\section{Optimization}
We have analyzed that the bottlenecks of Algorithm~\ref{algo:apriori_mining} 
lies in two factors. The size of each $Sr_s$ and the level the apriori runs.
In this section, we describe to optimization to reduce the two factors
respectively.

\subsection{Edge Simplification}
Each edge $e(s,t)$ in $Sr_s$ contains a time sequence $T$ 
which represent the co-occurrence of $s$ and $t$. We notice that the edge
between $s$ and $t$ is not always necessary. For example, if an edge has a
cardinality less than $K$, then it is not necessary to include this edge to $Sr_s$.
This motivates to simplify the edges in $Sr_s$ to boost the overall performance.

We first define a \emph{partly candidate} sequence as follows:
\begin{definition}[Partly Candidate Sequence]
Given a set of pattern parameters $M,L,K,G$, a sequence $T$ is 
a \emph{partly candidate} sequence if and only if exists a interval $(i,j)$
such that $T'=T[i:j]$ is a sequence conforming to $L,K,G$.
\end{definition}

For example, let $L = 2, K = 4, G = 2$, sequence $T_1=(1,2,4,5,9,10,11)$ 
is a \emph{partly candidate sequence} since $T_1[0:3] = (1,2,4,5)$ is a valid
pattern wrt. $L,K,G$. In contrast, $T_2=(1,2,5,6,7)$ is not a valid pattern.

Observing that only partly candidate sequence can be potential included in a 
pattern. Therefore, given an edge $e(s,t)=T \in Sr_s$, if $T$ is not a partly
candidate sequence, then this edge can be pruned from $Sr_s$. To efficiently
test whether a given sequence is partly candidate, we define the \emph{pseudo-consecutiveness} 
of a sequence as follows:

\begin{definition}[Pseudo-consecutiveness]
Given a parameter $G$, a sequence $T$ is \emph{pseudo-consecutive} if and only
if for any $i\in (0, |T|)$, $T[i] - T[i-1] \leq G$
\end{definition}

For example, let $G = 2$, then the sequence $T_1=\{1, 2, 4, 5\}$ is pseudo-consecutive
while $T_2=\{1,4,5,6,8\}$ is not pseudo-consecutive.

Based on the pseudo-consecutiveness, our edge simplification and prune process takes
two scan of an input $T$ as shown in Algorithm~\ref{algo:simp_prune}. In the 
first round of scan, the consecutive portion of $T$ with size less than $L$ is removed.
In the second round of scan, the pseudo-consecutive portion of $T$ with size less than $K$
is removed. Clearly the simplification algorithm runs in $O(|T|)$ time.

\begin{algorithm}
\caption{Edge Simplification}
\label{algo:simp_prune}
\begin{algorithmic}
\Require $T$
\State{Remove the consecutive portion with size less than $L$}
\State $c \gets 0$
\For {$i \in (0,...,|T|)$}
	\If{$T[i] - T[i-1] != 1$} 
		\If{$i - c < L$} 
			\State $T$ remove $[c:i)$
		\EndIf
		\State $c \gets i$
	\EndIf
\EndFor
\State{Remove the pseduo-consecutive portion with size less than $K$}
\State $s\gets 1$, $c\gets 0$
\For{$i \in (0: |T|)$}
	\If{$T[i] - T[i-1] > G$}
		\If{$s < K$}
			\State $T$ remove $[c:i)$
		\EndIf
		\State $c \gets i$
		\State $s \gets 1$
	\Else
		\State $s++$
	\EndIf
\EndFor
\end{algorithmic}
\end{algorithm}