### 2016-04-14 system.tex

\section{Preliminary on Apache Spark}
Apache Spark is the modern parallel processing platform
based on Resilient-Distributed-Dataset (RDD), which
is open-sourced by UC Berkeley. Spark differs from
traditional MapReduce framework (e.g., Hadoop) in the 
following features:
\begin{enumerate}
\item{It supports general computation DAGs beyond the two-stage MapReduce topologcy.}
\item{The execution engine can tolerate the loss of any works}
\item{It provides an in-memory storage abstraction called Resilient Distributed Datasets (RDDs). that 
lets applications keep data in memory, and automatically reconstructs the lost partitions upon failures.}
\end{enumerate}

In MapReduce-like systems, the major bottleneck is the \emph{shuffle} operation. 
In the \emph{shuffle} operation, data are re-partitioned and send to executors
via networks. Such operation is more costly than CPU and disk I/Os, which contribute
most in MapReduce-like applications.

TO BE ADDED....

\section{Mining Generalized Co-moving Pattern}
The outline of GCMP mining is to first cluster objects at
each snapshots and followed by mining the patterns among clusters
of different snapshots. In Spark scenario, 
the first step can be easily performed in parallel. 
This is because the objects at each snapshots are independent. Therefore, we
can partition the trajectories based on snapshots to achieve parallelism. 
In contrast, naively design the 
the second step may suffer from inter-executor communications.

\subsection{Temporal Replication and Mining}
The straightforward method of parallelizing GCMP mining is to partition
the trajectories in the temporal domain. A simple but effective method
is to group neighborhood snapshots in to a partition, such that all patterns
can be mined within a partition. In order to achieve the \emph{completeness},
some of the snapshots need to be replicated on multiple executors.
We thus call this method the \emph{Temporal Replication and Mining} approach. 
The algorithm is presented as in Algorithm~\ref{algo:trm_overview}.

\begin{algorithm}
\caption{Temporal Replication and Mining}
\label{algo:trm_overview}
\begin{algorithmic}
\Require list of $\langle t, S_t \rangle$ pairs
\State {Map Phase}
\ForAll{$\langle t, S_t \rangle$}
	\ForAll{$i \in 1...(K-1)*G+K$}
		\State emit a $\langle t-i, S_t \rangle$ pair
	\EndFor 
\EndFor
\State {Shuffle Phase}
\ForAll{$\langle t, S \rangle$ pair} 
\State group-by $t$, emit a $\langle t, Par_t\rangle$,
\State  where $Par_t = \{S_t, S_{t+1}, .. S_{t+(K-1)*G+K}\} $
\EndFor
\State {Reduce Phase}
\ForAll{$\langle t,Par_t \rangle$}
\State minePattern($Par_t$)
\EndFor

\end{algorithmic}
\end{algorithm}

\subsubsection{Temporal Replication Partition}
In order to reduce the shuffling cost, the data to be replicated should be
kept to a minimum. To control the number of snapshots that are replicated, 
we use the following theorem to facilitate a partition:

\begin{theorem}[Completeness of Replication Partition]
\label{thm:replication_partition}
For each snapshot $S_t$, a partition $Par_t = \{S_t, S_{t+1},...,S_{t+(K-1)*G+K}\}$ 
is created. Then, for any global pattern $P$, there must exist at least one partition $Par_i$
such that $P$ is also a pattern in $Par_i$.
\end{theorem}
\begin{proof}
PROOF BY CONTRADICTION, OMITTED FOR NOW
\end{proof}

Utilizing Theorem~\ref{thm:replication_partition}, we create, for each snapshot $S_t$, 
a partition containing its next $(K-1)*G+K$ snapshots. Each partition is then
sending to the executors as a task. Since any global pattern must exists in 
one of the partitions, we can mine the patterns from each partitions independently.

\subsubsection{Temporal Replication Mining}
After replication, each task in Spark will process an interval of snapshots $Par_i$. The next
step is to mine the GCMPs from $Par_i$. We notice that, for each 
partition, we only need to find the patterns that
are contained in the first snapshots. Therefore we design a line-sweep method for mining 
such patterns. DETAILS OMITTED FOR NOW.

The Temporal Replication and Mining approach though achieves parallelism from independent
partitions, it requires to replicate the data multiple times. 
Specifically, each snapshots are copied $(K-1)*G+K$ times. In swarm
case, $G$ is as large as $|T|$. Then, it is equivalent to shuffle entire dataset 
to each executor, which is clearly inefficient.

\subsection{Star Partition and Mining}
To develop a method that achieves parallelism under any pattern parameters, 
we propose a solution called \emph{Star Partition and Mining}. In SPM,
we design a novel object-based partition method named star partition. In star partition,
instead of partitioning trajectories in temporal domain, we partition trajectories 
in object domain. After partitioning, we use the \emph{Apriori}-like 
method to mine the patterns out of each partition. 
The outline of the start partition and mining is as in Algorithm~\ref{algo:spm_overview}

\begin{algorithm}
\caption{Star Partition and Mining}
\label{algo:spm_overview}
\begin{algorithmic}
\Require list of $\langle t, S_t \rangle$ pairs
\State {Map phase}
\ForAll{$C \in S_t$}
	\ForAll {$(o_1 ,o_2) \in C \times C$}
		\State emit a $\langle o_1, o_2, \{t\}\rangle$ triplet
	\EndFor
\EndFor

\State {Shuffle phase}
\ForAll{$\langle o_1, o_2, \{t\}\rangle$ triplets} 
	\State group-by $o_1$, emit $\langle o_1, Sr_{o_1} \rangle$ 
	\State group-by $o_2$, emit $\langle o_2, Sr_{o_2} \rangle$
\EndFor

\State {Reduce phase}
\ForAll{$\langle o, Sr_{o} \rangle$}
\State Apriori($Sr_o$)
\EndFor

\end{algorithmic}
\end{algorithm}

\subsubsection{Star Partition}
The purpose of star partition is to find the group of object that 
could potentially form a pattern. In order to achieve parallelism, 
we find and group, for each object $o$, all other objects that connect to $o$ in some
snapshots. By so doing, the patterns containing $o$ can be discovered within
each groups. Technically, we create a connection graph to represent the
connectivity among objects. A connection graph is an undirected graph $G=(V:E)$, where 
each $v \in V$ represents an object. An edge $e(s,t)= T$ 
represents the connection between vertex $s,t$ at all times in $T$, 
i.e. $\forall t \in T, C_t(s) = C_t(t)$. 
Given a vertex $s$, the star of $s$, $Sr_s$ is the set of incidental edges of $s$ in $G$.
In star partition, the trajectories are partitioned based on objects and their stars. Then
each partition is send to an executor as a task.
We state the completeness of star-partition as in the following theorem:
\begin{theorem}[Completeness of Star Partition]
Given a connection graph $G$, for any global pattern $P$, there exists a $s \in G.V$ such 
that $P$ is a pattern in $Sr_s$.
\end{theorem}

\begin{proof}
OMITTED FOR NOW
\end{proof}

Based on the above theorem, each task can be processed independently by executors. 
Thus, the inter-executor communication is avoided during the mining phase. It is notable that, the 
replication of data is $O(|\mathbb{O}|^2|T|)$, which is regardless the parameters of patterns.
In later sections, we will describe optimization techniques to reduce the replications.

\subsubsection{Apriori Mining}
For each task, in the mining phase, we need to find the patterns that conform to the parameter
setting. To systematically discover the patterns, we design the \emph{Apriori Mining} method which
is similar to frequent item mining. If a candidate pattern's object set is of size $R$, we call
this candidate a $R$-pattern. Our apriori mining algorithm tries to generate all $R$-pattern in
each iteration. As a start, the $2$-pattern is generated from edges in $Sr_s$. In particular,
for each $e(s,t)=T$, pattern $p=(\{s,t\}, T)$ is formed. During each iteration $R$, 
we generate $R+2$-cluster pattern by joining $(R+1)$-cluster patterns with $2$-cluster patterns. 
The join between $p_1=(O_1, T_1)$ and $p_2=(O_2, T_2)$ would generate a new pattern $p_3=(O_1 \cup O_2, T_1 \cap T_2)$.
Our mining algorithm stops where no further patterns are generated. The algorithm is illustrated as in Algorithm~\ref{algo:apriori_mining}.

\begin{algorithm}
\caption{Apriori Mining}
\label{algo:apriori_mining}
\begin{algorithmic}
\Require{$Sr_s$}
\State { Lv $\gets \{\}$}
\State { Lv2 $\gets \{\}$}
\ForAll{$e(s,t) = T \in Sr_s$}
\State Lv2.add($\langle \{s,t\}, T \rangle$);
\State Lv $\gets$ Lv2;
\EndFor
\While{true}
	\If{Lv is not empty} 
		\State{LvCand $\gets \{\}$ }
		\ForAll{$cand_v \in Lv$}
			\ForAll{$cand \in Lv2$}
				\State $p \gets cand_v$ join $cand$
				\If{$p.T$ is partly valid} 
					\State LvCand.add($p$)
				\EndIf
			\EndFor
		\EndFor
		\State {Lv $\gets$ LvCand}
	\Else
		\State{break}
	\EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{algo:apriori_mining} takes exponential complexity to mine GCMP. There
are two major factors dragging the performance. First, the size of $Sr_s$ affects
the number initial size of $2$-patterns. Second, the level that apriori runs. In later
sections, we exploit the property of GCMP to reduce the two factors.