### 2016-04-14 optimization.tex
Example and Proofs needed

\section{Optimization}
We have analyzed that the bottlenecks of Algorithm~\ref{algo:apriori_mining} 
lies in two factors. The size of each $Sr_s$ and the level the apriori runs.
In this section, we describe to optimization to reduce the two factors
respectively.

\subsection{Edge Simplification}
Each edge $e(s,t)$ in $Sr_s$ contains a time sequence $T$ 
which represent the co-occurrence of $s$ and $t$. We notice that the edge
between $s$ and $t$ is not always necessary. For example, if an edge has a
cardinality less than $K$, then it is not necessary to include this edge to $Sr_s$.
This motivates to simplify the edges in $Sr_s$ to boost the overall performance.

We first define the \emph{Pseudo-consecutiveness} of a time sequence as follows:
\begin{definition}[Pseudo-consecutiveness]
Given a parameter $G$, a sequence $T$ is \emph{pseudo-consecutive} if and only
if for any $i\in (0, |T|)$, $T[i] - T[i-1] \leq G$
\end{definition}
For example, let $G = 2$, then the sequence $T_1=\{1, 2, 4, 5\}$ is pseudo-consecutive
while $T_2=\{1,4,5,6,8\}$ is not pseudo-consecutive.

A pseudo-consecutive subsequence of $T$ is a subsequence $T' \subseteq T$
and $T'$ is pseudo-consecutive. A maximal pseudo-consecutive subsequence 
$T^m$ of $T$ is a pseudo-consecutive subsequence of $T$ and there is 
no superset of $T^m$ that is also a pseudo-consecutive subsequence of $T$.
Clearly, a sequence $T$ can be decomposed into several maximally pseudo-consecutive 
subsequences. We then define a \emph{partly candidate} sequence as follows:

\begin{definition}[Partly Candidate Sequence]
Given the pattern parameters: $L,K,G$, a sequence $T$ is 
a \emph{partly candidate} sequence if exist one of its maximal pseudo-consecutive
subsequence $T'$ such that $T'$ confirms to $L,K,G$.
\end{definition}

For example, let $L = 2, K = 4, G = 2$, sequence $T_1=(1,2,4,5,6,9,10,11)$ 
is a \emph{partly candidate sequence} since $T_1[0:3] = (1,2,4,5,6)$ is a valid
pattern wrt. $L,K,G$. In contrast, $T_2=(1,2,5,6,7)$ is not a valid partly candidate sequence.

Observing that only partly candidate sequence can be potentially included in a 
pattern. Therefore, given an edge $e(s,t)=T \in Sr_s$, if $T$ is not a partly
candidate sequence, then this edge can be pruned from $Sr_s$. To efficiently
test whether a given sequence is partly candidate, we define the \emph{Fully Candidate Sequence}:

\begin{definition}[Fully Candidate Sequence]
Given the pattern parameters: $L,K,G$, a sequence $T$ is a \emph{Fully Candidate} 
if and only if for any of its maximal pseudo-consecutive sequence $T'$, $T'$ 
is partly consecutive.
\end{definition}

For example, let $L = 2, K = 4, G = 2$, sequence $T_1=(1,2,4,5,6,9,10,11)$ is 
not a fully candidate sequence since one of its maximal pseudo-consecutive sequence $(9,10,11)$
is not a partly candidate sequence. In contrast, sequence $T_2=(1,2,4,5,6)$ is 
a fully candidate sequence.

Based on the fully candidate sequence, we can reduce an sequence $T$ to a 
fully candidate sequence by striping out its non-partly candidate maximal pseudo-consecutive 
sequences. The reduction works as in Algorithm~\ref{algo:simp_prune}. It takes two
rounds of scan of an input $T$. In the first round of scan,
the consecutive portion of $T$ with size less than $L$ is removed.
In the second round of scan, the pseudo-consecutive portion of $T$ with size less than $K$
is removed. Clearly the simplification algorithm runs in $O(|T|)$ time.

\begin{algorithm}
\caption{Edge Simplification}
\label{algo:simp_prune}
\begin{algorithmic}
\Require $T$
\State{Remove the consecutive portion with size less than $L$}
\State $c \gets 0$
\For {$i \in (0,...,|T|)$}
	\If{$T[i] - T[i-1] != 1$} 
		\If{$i - c < L$} 
			\State $T$ remove $[c:i)$
		\EndIf
		\State $c \gets i$
	\EndIf
\EndFor
\State{Remove the pseduo-consecutive portion with size less than $K$}
\State $s\gets 1$, $c\gets 0$
\For{$i \in (0: |T|)$}
	\If{$T[i] - T[i-1] > G$}
		\If{$s < K$}   
			\State $T$ remove $[c:i)$
		\EndIf
		\State $c \gets i$
		\State $s \gets 1$
	\Else
		\State $s++$
	\EndIf
\EndFor
\end{algorithmic}
\end{algorithm}

We use the following theorem to state the completeness and correctness of our 
edge reduction algorithm.
\begin{theorem}[Completeness of Algorithm~\ref{algo:simp_prune}]
For a sequence $T$, let $T_1$ be the result of $T$ after algorithm~\ref{algo:simp_prune},
if $\exists T' \subseteq T$ and $T'$ is included in a pattern $P$, then $T' \subseteq T_1$.
\end{theorem}

\begin{proof}
OMITTED FOR NOW
\end{proof}
By utilizing the edge simplification technique, the size of $Sr_s$ can be greatly reduced. If
an edge cannot be reduced to a full candidate sequence, then it is directly removed from $Sr_s$.
If an edge can be reduced to a full candidate sequence, replacing itself with its full candidate sequence 
would result in a more compact storage.

\subsection{Pruning Apriori Mining}
During the apriori phase, we are repeatedly join candidate pattern in different levels to generate a larger set
of a pattern. We notice that, such joins can be early terminated 
by utilizing the property of GCMP. The intuition is that when we are trying to store a level of candidate, if its
temporal sequence is not a \emph{partly candidate} sequence, then it should able to generate new candidates and thus it
can be removed from that level of candidate. This \emph{apriori property} 
is explicitly described as in the follow theorem:

\begin{theorem}[Apriori Property of GCMP]
Given the temporal parameters $L,G,K$ of GCMP, for a candidate $cand$ in Algorithm~\ref{algo:apriori_mining},
if $cand.T$ cannot be reduced to a full candidate sequence, then $cand$ can be pruned.
\end{theorem}
\begin{proof}
OMITTED FOR NOW
\end{proof}

Utilizing the \emph{Apriori Property}, the new candidate in each level is greatly reduced, which is shown in
the experiment session.