### 2016-04-26 optimization.tex

\section{Optimization}
We have analyzed that the bottlenecks of Algorithm~\ref{algo:apriori_mining} 
lies in two factors. The size of each $Sr_s$ and the size of candidates in each level of Apriori.
In this section, we provide several optimizations to boost the bottlenecks.

\subsection{Edge Reduction by Direction}
The first spot for reducing the size of $Sr_s$ is to remove the replicated edges. As shown in Algorithm~\ref{algo:spm_overview}, each edge in the conceptual graph is replicated twice in generating the star-structure. The purpose of replication is to ensure the completeness of star partition. However, this replication can be avoided if we choose an appropriate way of partitioning.

We design the edge partitioning method by edge direction. Instead of building a conceptual graph that is undirected, we create the directed conceptual graph as follows: First, we assign each object a unique number. Then
for a cluster $C_t$ in snapshot $S_t$, for any pair $(u,v) \in C_t$, an edge $e(u,v) = \{t\}$ is created if $u < v$. It is easy to see that the directed conceptual graph is a DAG. We then create each $Sr_u$ by including all the outgoing edges of $u$. By so doing, each edge is assigned to only one star, thus avoids the replications. We use the following theorem to ensure the completeness of the edge direction method.

\begin{theorem}[Sound and Completeness of Edge Direction]
Star partition with edge direction is sound and complete.
\end{theorem}
\begin{proof}
It is notable that each star is a subset of original trajectories, thus the soundness is trivially true. For completeness, if $P$ is valid pattern, then let $s$ be the object of the smallest number in $P.O$, i.e., $s=\min_{o \in P.O}(o)$. Since $s$ is smallest and the all other objects in $P.O$ is connected with $s$. Therefore, $P.O \equiv Sr_s$, which indicates that $P.O$ is also a pattern in $Sr_s$.
\end{proof}
An example of edge direction is shown in Figure~\ref{fig:star_partition}. As shown,
by adapting the direction method, half the size of $S_r$'s is reduced. This clearly brings efficiency in both shuffling and apriori mining.

\subsection{Edge Simplification}
Each edge $e(s,t)$ in $Sr_s$ contains a time sequence $ET$ 
which represent the co-occurrence of $s$ and $t$. We notice that the edge between $s$ and $t$ is not always necessary. For example, if an edge has a
cardinality less than $K$, it is unnecessary to include this edge to $Sr_s$ since it cannot contribute to any patterns.
This motivates us to simplify the edges in $Sr_s$ to boost the overall performance.

We first define the \emph{Pseudo-consecutiveness} of a time sequence as follows:
\begin{definition}[Pseudo-consecutiveness]
Given a parameter $G$, a sequence $T$ is \emph{pseudo-consecutive} if and only
if for any $i\in [1, |T|]$, $T[i] - T[i-1] \leq G$
\end{definition}
For example, let $G = 2$, then the sequence $T_1=\{1, 2, 4, 5\}$ is pseudo-consecutive
while $T_2=\{1,4,5,6,8\}$ is not pseudo-consecutive.

A pseudo-consecutive subsequence of $T$ is a subsequence $T' \subseteq T$
and $T'$ is pseudo-consecutive. A maximal pseudo-consecutive subsequence 
$T^m$ of $T$ is a pseudo-consecutive subsequence of $T$ and there is 
no superset of $T^m$ that is also a pseudo-consecutive subsequence of $T$.
Clearly, a sequence $T$ can be decomposed into several maximally pseudo-consecutive 
subsequences. We then define a \emph{partly candidate} sequence as follows:

\begin{definition}[Partly Candidate Sequence]
Given the pattern parameters: $L,K,G$, a sequence $T$ is 
a \emph{partly candidate} sequence if exists one of its maximal pseudo-consecutive
subsequence $T'$ such that $T'$ confirms to $L,K,G$.
\end{definition}

For example, let $L = 2, K = 4, G = 2$, sequence $T_1=(1,2,4,5,6,9,10,11)$ 
is a \emph{partly candidate sequence} since $T_1[1:5] = (1,2,4,5,6)$ is a valid
pattern wrt. $L,K,G$. In contrast, $T_2=(1,2,5,6,7)$ is not a valid partly candidate sequence.

Observing that only partly candidate sequence can be potentially contribute to a 
pattern. Therefore, given an edge $e(s,t)=T \in Sr_s$, if $T$ is not a partly
candidate sequence, it can be pruned from $Sr_s$. To efficiently
test whether a given sequence is partly candidate, we define the \emph{Fully Candidate Sequence}:

\begin{definition}[Fully Candidate Sequence]
Given the pattern parameters: $L,K,G$, a sequence $T$ is a \emph{Fully Candidate} 
if and only if for any of its maximal pseudo-consecutive sequence $T'$, $T'$ 
is partly consecutive.
\end{definition}

For example, let $L = 2, K = 4, G = 2$, sequence $T_1=(1,2,4,5,6,9,10,11)$ is 
not a fully candidate sequence since one of its maximal pseudo-consecutive sequence $(9,10,11)$
is not a partly candidate sequence. In contrast, sequence $T_2=(1,2,4,5,6)$ is 
a fully candidate sequence.

Based on the fully candidate sequence, we can reduce an sequence $T$ to a 
fully candidate sequence by striping out its non-partly candidate maximal pseudo-consecutive 
sequences. The reduction works as in Algorithm~\ref{algo:simp_prune}. It takes two
rounds of scan of an input $T$. In the first round of scan,
the consecutive portion of $T$ with size less than $L$ is removed.
In the second round of scan, the pseudo-consecutive portion of $T$ with size less than $K$
is removed. Clearly the simplification algorithm runs in $O(|T|)$ time.

\begin{algorithm}
\caption{Edge Simplification}
\label{algo:simp_prune}
\begin{algorithmic}
\Require $T$
\State{Remove the consecutive portion with size less than $L$}
\State $c \gets 0$
\For {$i \in (0,...,|T|)$}
	\If{$T[i] - T[i-1] != 1$} 
		\If{$i - c < L$} 
			\State $T$ remove $[c:i)$
		\EndIf
		\State $c \gets i$
	\EndIf
\EndFor
\State{Remove the pseduo-consecutive portion with size less than $K$}
\State $s\gets 1$, $c\gets 0$
\For{$i \in (0: |T|)$}
	\If{$T[i] - T[i-1] > G$}
		\If{$s < K$}   
			\State $T$ remove $[c:i)$
		\EndIf
		\State $c \gets i$
		\State $s \gets 1$
	\Else
		\State $s++$
	\EndIf
\EndFor
\end{algorithmic}
\end{algorithm}

We use the following theorem to state the completeness and correctness of our 
edge reduction algorithm.
\begin{theorem}[Soundness and Completeness Edge Simplification]
Star partition with edge simplification is sound and complete.
\end{theorem}

\begin{proof}
Soundness of the star partition is not affected by edge simplification since each star is a subset of original trajectory. For completeness, consider a sequence $T$, let $T_1$ be the result of $T$ after first round of simplification. By the nature of the first round of simplification, if $P.T \subseteq T$, then $P.T \subseteq T_1$. In the second round of simplification, only the pseudoconsecutive parts with size less than $K$ are removed. This means, the removed parts is not able to contribute to any patterns. Thus if $P.T \subseteq T$, then $P.T \subseteq T'$. 
\end{proof}
By leveraging the edge simplification, the size of $Sr_s$ can be greatly reduced. If
an edge cannot be reduced to a full candidate sequence, then it is directly removed from $Sr_s$.
If an edge can be reduced to a full candidate sequence, replacing itself by the full candidate sequence results in a more compact storage.

\subsection{Pruning Apriori Mining}
During the apriori phase, we repeatedly join candidate patterns in different levels to generate a larger set
of a patterns. We notice that, such joins can be early terminated 
by utilizing the monotonic property of GCMP. The intuition is that when we are trying to store a candidate, if its
temporal sequence is not a \emph{partly candidate} sequence, then it cannot generate full candidate and thus it
can be pruned. This \emph{monotonic property} 
is explicitly described as in the follow theorem:

\begin{theorem}[Monotonic Property of GCMP]
Given the temporal parameters $L,G,K$, for a candidate $cand$ in Algorithm~\ref{algo:apriori_mining},
if $cand.T$ cannot be reduced to a partly candidate sequence, then $cand$ can be pruned.
\end{theorem}
\begin{proof}
Let $cand_i$ be a $i$-pattern. If $cand_i$ is generated from $cand_j$ where $j < i$, then if $cand_j.T \subseteq cand_i.T$. That is as the level goes up, the time sequence set of a candidate shrinks. Therefore, if $cand_j.T$ is not a partly candidate sequence, $cand_i.T$ cannot be a partly candidate sequence.
\end{proof}

With the help of the \emph{Monotonic Property}, the number of new candidates in each level is greatly reduced. We verify this in the experiment session as well.