### 2016-05-28 sec8_experiment.tex

\section{Experimental Study}
\label{sec:experiment}
\subsection{Implementation Issues}
We use Apache Spark~\footnote{http://spark.apache.org/} as the experimental
platform. Spark is one of the most popular MapReduce-like platform which
uses in-memory cache to gain high speedup against Apache Hadoop~\cite{}. Since
Spark directly supports the MapReduce paradigm, our algorithm is able to 
be easily implemented in Spark. In order to help reproduce our experiments,
we further address some implementation issues.
\subsubsection{Task Assignment}
In spark, each task in reduce phase is an Apriori mining phase of 
a star.
Although our star partition method is theoretically balanced, 
it is still necessary to assign equal number of tasks to each executors. 
Spark naturally uses hashing to partition data into tasks, where such
a partitioning does not care on the tasks size. In order to
fully utilize the clusters, it is important to perform a weight-aware
partition. In our implementation, we collect the number of edges
in each star after map phase. Afterwards, we use a simple \emph{best-fit} strategy,
where we assign stars in decreasing order with their sizes and each star
is assigned to the currently least-loaded executor.  HERE WE MAY HAVE ANOTHER BOUND
FROM LITERATURE, BUT I DIDN'T FIND YET. The injection of load balancing strategy
between map and reduce phase can be naturally implemented in Spark, where the 
map result can be cached and the reduce phase can be paused until when the
partition strategy is ready.

\subsubsection{Duplication Detection}
It is notable that the patterns discovered from different tasks (stars) could
be redundant due to containment relationship. For example, a pattern $\{a,b,c\}$
can be discovered from the star $Sr_a$, while the pattern $\{b,c\}$ can be discovered
from the star $Sr_b$. Though in most applications, such a duplicate pattern
is permitted, we offer an option to eliminate these patterns. The strategy is 
to broadcast each reducers output to every other reducers. This can be
efficiently done via \emph{broadcast} variables~\footnote{http://spark.apache.org/docs/latest/programming-guide.html\#broadcast-variables} in Spark. Afterwards, each reduce can check 
whether any resulted patterns are subsumed and thus filter those patterns.
Theoretically, advanced techniques, such as Bloom Filters, can be applied to efficiently
deal with the duplication detection. However, as the number of final patterns are
normally quit small, we leave the exploration for those techniques to the future.

\subsubsection{Handling Overlapping Clusters}
When handling patterns such as \emph{flock} and \emph{group}, disk-based clustering
on objects are applied. Such a clustering method may result one object belonging to
multiple clusters. In such a case, just keeping the timestamps in the edge
of connection graph is insufficient. Instead, we extend every timestemp $t$
to a pair $\langle t,C \rangle$, where $C$ is the set of clusters objects
belong to at time $t$. The only adaption we need to take the join during
apriori phase. Given two timestamp set $T_1$ and $T_2$, the join result of
$T_1$ and $T_2$ instead of being $\{\forall t | t\in T_1 \wedge t_\in T_2\}$,
it changes to $\{\forall (t,C) | t\in T_1 \wedge t \in T_2 \wedge C = (T_1.C \cap T_2.C) \wedge C \neq \emptyset\}$.
It is obvious to see the \emph{edge simplification} and \emph{candidate pruning} 
still holds under this new setting.

\subsection{Experimental Setup}
Our experiments run on a 9-node cluster, with Apache Yarn as
the cluster manager. We use 1 node for Yarn resource manager, 
and use the remaining 8 nodes as executors in Spark. In the 
cluster, each node is uniformly equipped with a 2.2GHz quad-core CPU
with 32 GB memory. Inter-node communication is carried by 
the 1Gbps Ethernet.  Some critical configuration of Spark is 
as follows:
\begin{table} [h]
\centering
\begin{tabular}{|l|c|}
\hline 
Parameter & Value  \\ 
\hline 
Java Version & 1.7.0 \\ 
\hline 
spark.driver.memory & 2GB \\ 
\hline 
spark.executor.cores & 2  \\ 
\hline 
spark.executor.instances & 11 \\ 
\hline 
spark.executor.memory & 7GB \\ 
\hline 
spark.master & yarn-cluster \\ 
\hline 
spark.serializer & KryoSerializer \\ 
\hline 
\end{tabular} 
\end{table}

We setup HDFS on the same cluster and all trajectory data
are initially stored in the HDFS. We prepare two set of
large-scaled trajectories with the following properties: