### 2016-06-25 sec7_conclusion.tex
A new draft
Flighting

\section{Conclusion and Future Work}
\label{sec:concl}