### 2016-06-25 sec7_conclusion.tex
Need to add three more experiments
Three more charts

\section{Conclusion}
\label{sec:concl}
In this paper, we present our framework SPARE for detecting General Co-Movement Patterns (GCMP) over large-scale trajectories. Our GCMP is versatile in representing all existing co-movement patterns while being free from the \emph{loose-connection} anomaly. Our SPARE framework, comprising a novel star-partition with several optimization techniques, has shown superior performances and scalability in three real datasets with hundreds of millions trajectories.