### 2016-06-26 sec6_experiment.tex

\section{Experimental Study}
\label{sec:exp}
In this section, we evaluate the efficiency and scalability of our proposed distributed GCMP detectors on real trajectory datasets. All the experiments are carried out in a cluster with $12$ nodes, each equipped with four quad-core $2.2$GHz Intel processors, $32$GB memory and gigabit Ethernet.

\textbf{Environment Setup}: We use Yarn\footnote{\url{http://hadoop.apache.org/docs/current/hadoop-yarn/hadoop-yarn-site/YARN.html}} to manage our cluster. We pick one machine as the Yarn's master node, and the remainings reserve one core and $2$GB memory for Yarn processes. We deploy our GCMP detector on Apache Spark 1.5.5~\cite{zaharia2012resilient} with the remaining $11$ nodes as the computing nodes.
To fully utilize the computing resources, we configure each node to run five executors, each taking three cores and $5$GB memory. In Spark, one of the $55$ executor is taken as the Application Master for coordination, therefore our setting results in $54$ executors.
All our implementations as well as cluster setups are open-sourced in Github\footnote{\url{https://github.com/fanqi1909/TrajectoryMining/}}.

\textbf{Datasets}: We use three real trajectory datasets in different application scenarios:
\begin{itemize}
\item{Shopping}\footnote{\url{http://www.irc.atr.jp/crest2010_HRI/ATC_dataset/}}. The dataset contains
  trajectories of visitors in the ATC shopping center in Osaka. To better capture the indoor activities, the visitor locations are sampled every half second, resulting in $13,183$ long trajectories. 
\item{GeoLife}~\footnote{\url{http://research.microsoft.com/en-us/projects/geolife/}}. The dataset essentially keeps all the travel records of 182 users for a period
of over three years, including multiple kinds of transportation modes (walking, driving and taking public
transportation). For each user, the GPS information is collected periodically and 91 percent of the trajectories
are sampled every 1 to 5 seconds.
\item{Taxi}. The dataset tracks the trajectories of $15,054$ taxies in Singapore over August 2008. The sampling
rate is around 30 seconds. 
\end{itemize}

\textbf{Preprocessing}: We replace timestamps with global sequences (starting from $1$) for each dataset. 
We set a fixed sampling rate for each dataset (i.e., Geolife = 5 seconds, Shopping=0.5 seconds, Taxi = 30 seconds)
and use linear interpolation to fill missing values.
%For each dataset, we set a fixed sampling rate (CLEARLY STATE THE SAMPLING RATE FOR EACH DATASET) and use linear interpolation to fill missing values. 
%Finally, we obtain a set of trajectories with equal length. 
%Finally, we obtain a set of dense trajectories with equal length. 
For the clustering method, we use DBSCAN~\cite{ester1996density} and customize its two parameters $\epsilon$ (proximity threshold) and $minPt$ (the minimum number of points required to form a dense region). We set $\epsilon=5$, $minPt=10$ for GeoLife and Shopping datasets; and $\epsilon=20$, $minPt=10$ for Taxi dataset. Note that other clustering methods or settings can also be applied. 
%The clustered snapshots are stored in HDFS as $\langle t, S_t \rangle$ pair, where $t$ is the timestamp, $S_t$ contains the clusters at snapshot $t$. 
After preprocessing, the statistics of the three datasets are listed in Table~\ref{exp:dataset}.

\begin{table} [h]
\center
\small
\begin{tabular}{|l|l|l|l|}
\hline
 \textbf{Attributes}& \textbf{Shopping} &  \textbf{GeoLife} &  \textbf{SingTaxi} \\ 
\hline 
\# objects  & 13,183 & 18,670 & 15,054\\ 
\hline
%\# average ts & 3,114  & 2,924 & 19,667 \\ 
%\hline
\# data points  & 41,052,242 & 54,594,696 & 296,075,837\\ 
\hline
\# snapshots  & 16,931 & 10,699 & 44,364\\ 
\hline
\# clusters  & 211,403  & 206,704& 536,804\\
\hline
avg. cluster size  & 171 & 223 & 484\\
\hline
\end{tabular}
\caption{Statistics of data set}
\label{exp:dataset}
\end{table}

\textbf{Parameters}: To systematically study the performance of
our algorithms, we conduct experiments on various parameter settings. The parameters to be evaluated are listed in Table~\ref{tbl:parameters}, with default settings in bold. 
\begin{table}[h]
\small
\begin{tabular}{c|l|l}
\hline 
\textbf{Variables} & \textbf{Meaning} & \textbf{Values} \\ 
\hline 
M & min size of object set &  5, 10,  \textbf{15}, 20, 25 \\ 
\hline 
K & min duration & 120, 150, \textbf{180}, 210, 240 \\ 
\hline 
L & min local duration & 10, 20, \textbf{30}, 40,50 \\ 
\hline 
G & max gap & 10, 15, \textbf{20}, 25, 30 \\ 
\hline
$O_r$ & ratio of objects & 20\%,40\%, 60\%, 80\%, \textbf{100\%} \\ \hline
$T_r$ & ratio of snapshots & 20\%,40\%, 60\%, 80\%, \textbf{100\%} \\ \hline
N & number of machines & 1, 3, 5, 7, 9, \textbf{11}\\ 
\hline 
\end{tabular} 
\caption{Variables and their default values}
\label{tbl:parameters}
\end{table}

\subsection{Performance Evaluation}
\begin{figure*}[t]
\centering
    \begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/shopping_vary_M.eps}
        \caption{Shopping vary $M$}
    \end{subfigure}
    \begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/shopping_vary_K.eps}
        \caption{Shopping vary $K$}
    \end{subfigure}
    \begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/shopping_vary_L.eps}
        \caption{Shopping vary $L$}
    \end{subfigure}
       \begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/shopping_vary_G.eps}
        \caption{Shopping vary $G$}
    \end{subfigure}

	\begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_M.eps}
        \caption{Geolife vary $M$}
    \end{subfigure}
    \begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_K.eps}
        \caption{Geolife vary $K$}
    \end{subfigure}
    \begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_L.eps}
        \caption{Geolife vary $L$}
    \end{subfigure}
       \begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_G.eps}
        \caption{Geolife vary $G$}
    \end{subfigure}

    \begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/taxi_vary_M.eps}
        \caption{Taxi vary $M$}
    \end{subfigure}
    \begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/taxi_vary_K.eps}
        \caption{Taxi vary $K$}
    \end{subfigure}
    \begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/taxi_vary_L.eps}
        \caption{Taxi vary $L$}
    \end{subfigure}
       \begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/taxi_vary_G.eps}
        \caption{Taxi vary $G$}
    \end{subfigure}       
\caption{Performance of SPARE and TRPM on real datasets under different pattern parameters.}
\label{exp:performance_vary}
\end{figure*}
\textbf{Varying $M$}: Figures~\ref{exp:performance_vary} (a),(e),(i)
present the performance with increasing $M$. The SPARE framework demonstrates a clear superiority over TRPM framework, with a boosting factor of  $2.7$ times in Shopping, $3.1$ times in Geolife and
$7$ times in Taxi. As $M$ increases, the running time of both frameworks slightly improves because the number of clusters in each snapshot drops, generating fewer valid candidates.

\textbf{Varying $K$}: The performance with increasing $K$ is shown in Figure~\ref{exp:performance_vary} (b),(f),(j).  SPARE tends to run faster, whereas the performance of TRPM degrades dramatically. This is caused by the \emph{sequence simplification} procedure in SPARE, which can prune many candidates with large $K$. However, the line sweep algorithm in TRPM does not utilize such property for pruning. It takes longer time because more replication data has to be handled in each partition.

\textbf{Varying $L$}: Figures~\ref{exp:performance_vary} (c)(g)(k) present the performances with varying $L$. When $L=10$, SPARE can outperform TRPM by around $10$ times. We also observe that there is a significant performance improvement for TPRM when $L$ increases from $10$ to $20$ and later the running time drops smoothly. 
This is because $\eta$ is proportional to $O(K*G/L+L)$. When $L$ is small (i.e., from $10$ to $20$),
$\eta$ decreases drastically. As $L$ further grows, $\eta$ varies less significantly.

\textbf{Varying $G$}: Figures~\ref{exp:performance_vary} (d)(h)(l) present the performances with increasing $G$.  TRPM is rather sensitive to $G$. When $G$ is relaxed to larger values, more valid patterns would generate. TPRM has to set a higher replication factor and its running time degrades drastically when $G$ increases from $20$ to $30$. In contrast, with much more effective pruning strategy, our SPARE scales very well with $G$.

\textbf{Varying $O_r$}: Figures~\ref{exp:performance_vary_OT} (a)(b)(c) present 
the performances with increasing number of moving objects. Both TRPM and SPARE take longer time to find patterns in a growing database. We can see that the performance gap between SPARE and TRPM is widened as more objects are involved, which shows SPARE is more scalable.

\textbf{Varying $T_r$}: Figures~\ref{exp:performance_vary_OT} (d)(e)(f) present 
the performances with increasing number of snapshots. As $T_r$ increases, SPARE scales much better than TRPM due to its effective pruning in the temporal dimension.

\begin{figure}[h]
\centering
	\begin{subfigure}[b]{0.22\textwidth}
	 \includegraphics[width=\textwidth]{/exp/performance/shopping_vary_o.eps}
        \caption{Shopping vary $O_r$}
    \end{subfigure}
 	 \begin{subfigure}[b]{0.22\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_o.eps}
        \caption{Geolife vary $O_r$}
    \end{subfigure}
    	 \begin{subfigure}[b]{0.22\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/taxi_vary_o.eps}
        \caption{Taxi vary $O_r$}
    \end{subfigure}
    \begin{subfigure}[b]{0.22\textwidth}
	 \includegraphics[width=\textwidth]{/exp/performance/shopping_vary_t.eps}
        \caption{Shopping vary $T_r$}
    \end{subfigure}
 	 \begin{subfigure}[b]{0.22\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_t.eps}
        \caption{Geolife vary $T_r$}
    \end{subfigure}
    	 \begin{subfigure}[b]{0.22\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/taxi_vary_t.eps}
        \caption{Taxi vary $T_r$}
    \end{subfigure}
 \caption{Scalability of SPARE and TRPM wrt. $O_r$ and $T_r$}
 \label{exp:performance_vary_OT}
\end{figure}

\subsection{Analysis of SPARE framework}
In this part, we extensively evaluate the advantage brought by the
sequence simplification and load balancing techniques.

\subsubsection{Power of sequence simplification}
To study the power of sequence simplification,
we collect two types of statistics: (1) the number of pairs that
are shuffled to the reducers and (2) the number of pairs that
are fed to the Apirori enumeration. The difference between these two values is the number of size-$2$ candidates pruned by the sequence simplification.
The results in Table~\ref{tbl:pruning} show that the \emph{sequence simplification} is very powerful and eliminates nearly 90 percent of the object pairs, which significantly reduces the overhead of subsequent Apriori enumerator.

\begin{table}[h]
\centering
\begin{tabular}{|l|c|c|c|}
\hline 
\textbf{Data Set} & \textbf{Shopping} & \textbf{GeoLife} & \textbf{Taxi} \\ 
\hline 
Before pruning & 878,309 &  1,134,228 & 2,210,101 \\ 
\hline 
After pruning & 76,672 & 123,410 & 270,921 \\ 
\hline 
Prune ratio & 91.2\% & 89.1\% & 87.7\% \\ 
\hline 
\end{tabular}

\caption{Pruning power of SPARE}
\label{tbl:pruning}
\end{table}

\begin{figure}[h]
\centering
	  \includegraphics[width=0.3\textwidth]{/exp/spare/detail.eps}
    \caption{Break-down cost of SPARE and SPARE-RD}
    \label{exp:wl}
%	\begin{subfigure}[b]{0.3\textwidth}
%        \includegraphics[width=\textwidth]{/exp/spare/detail.eps}
%        \caption{Decomposed running time}
%    \end{subfigure}
%    
% 	\begin{subfigure}[b]{0.22\textwidth}
%       \includegraphics[width=\textwidth]{/exp/spare/wl-shopping.eps}
%       \caption{Workload in Shopping}
%   \end{subfigure}
%    \begin{subfigure}[b]{0.22\textwidth}
%        \includegraphics[width=\textwidth]{/exp/spare/wl-geolife.eps}
%        \caption{Workload in Geolife}
%    \end{subfigure}  
%    \begin{subfigure}[b]{0.22\textwidth}
%        \includegraphics[width=\textwidth]{/exp/spare/wl-taxi.eps}
%        \caption{Workload in Taxi}
%    \end{subfigure}
%\caption{Load balance of SPARE and SPARE-LB}
\end{figure}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|l|l|l|}
\hline
\multirow{2}{*}{\textbf{Dataset}} & \multicolumn{2}{l|}{\textbf{SPARE-RD}} & \multicolumn{2}{l|}{\textbf{SPARE}} \\ \cline{2-5} 
                         & Straggler        & STD        & Straggler       & STD      \\ \hline
Shopping                 & 295              & 41         & 237             & 21       \\ \hline
GeoLife                  & 484              & 108        & 341             & 56       \\ \hline
Taxi                     & 681              & 147        & 580             & 96       \\ \hline
\end{tabular}
  \caption{Statistics of execution time among executors,units are in seconds.}
  \label{tbl:strags}
\end{table}

\subsubsection{Load Balance}
To study the effect of load balance in the SPARE framework, we use random task allocation (the default setting of Spark) as a baseline, denoted by SPARE-RD, and compare it with our best-fit method. In best-fit, the largest unassigned star is allocated to the currently most empty reducers.
Figure~\ref{exp:wl}(a) shows the break-down costs in the map-reduce stages for SPARE and SPARE-RD. We observe that the map and shuffle time of SPARE and SPARE-RD are identical. The difference is that SPARE spends additional time cost to make an allocation plan for load balance (around $4\%$ of the total cost), resulting in significant savings in the reduce stage (around $20\%$ of the total cost). We also report the cost of straggler, i.e., the longest job, in Table~\ref{tbl:strags}, whose results clearly verify the effectiveness of our allocation strategy for load balance.

\subsection{Parallelism}
Finally, we examine the parallelism of our proposed frameworks in both multi-core and multi-node environments. First, we use single-node mode and examine the performance with multi-cores. We also compare it with the  state-of-the-art solutions for swarm and platoon patterns since they were designed  to run in a centralized sever. Second, we examine the pattern detection time when more and more computing resources (number of machines) are available. The results are reported in Figure~\ref{exp:scalability}. First, the centralized schemes are not suitable to discover patterns in large-scale trajectory databases. It takes nearly $16$ hours to detect \emph{platoon}s in Taxi data in a single machine. In contrast, when utilizing the multi-core environment, our TRPM and SPARE achieves 2.4 times and 7.1 times speedup respectively. Second, we see that both TRPM and SPARE demonstrate promising scalability in terms of machines available. The running time almost decreases linearly as more machines are used. When there are $11$ nodes in the cluster, the performance of SPARE is $94$ times better than the state-of-the-art \emph{platoon} detector in a centralized server.

\begin{figure*}[t]
\centering
\begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{/exp/spare/scalability-shopping.eps}
        \caption{Shopping vary $N$}
    \end{subfigure}
 	 \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{/exp/spare/scalability-geolife.eps}
        \caption{Geolife vary $N$}
    \end{subfigure}
    	 \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{/exp/spare/scalability-taxi.eps}
        \caption{Taxi vary $N$}
    \end{subfigure}
 \caption{Comparisons among TRMP, SPARE and Platoon}
 \label{exp:scalability}
\end{figure*}