### 2016-06-26 sec7_conclusion.tex

\section{Conclusions and Future Work}
\label{sec:concl}
In this paper, we proposed a generalized co-movement pattern to unify those proposed in the past literature. We also devised two types of parallel frameworks in Spark that can scale to support pattern detection in a trajectory database with hundreds of millions of points. The efficiency and scalability were verified by extensive experiments on three real datasets. In the future work, we are interested to examine co-movement pattern detection in streaming data for real-time monitoring. How to extend the current parallel framework to support other types of advanced patterns is also of interest.