### 2016-08-26 ReviewComments/meta.tex
Revision
almost completed

\section{Response to Meta Reviewer}

\emph{Novelty: The GCMP generalization is not particularly novel. Please elaborate on the novelty of the work.}

\response{In addition to the points in Response 2, another novelty of our work lies in the 
techniques. 
Although the idea of star-partition and apriori-partition are well-known, 
linking them together to solve the problem in trajectory domain has not been attempted before.
Besides, we input heavy details on these algorithms (e.g., theoretical bounds of partition and
anti-monotonicity) which have not been proposed before.
}

\emph{Parameter setting: Are we really interested in all sets of movements beyond a cardinality of size M? Please elaborate
on the motivation.}

\response {Similar responses with Response 3. }

\emph{Spark implementation: the paper references Spark as the implementation platform but the algorithm is limited to MapReduce. Spark has capabilities beyond MapReduce such as window functions. Please provide an implementation that uses the relevant features of Spark, or a convincing discussion of how these features can be useful and why they were not used.}

\response{Similar responses with Response 1 and 5. }

\emph{More details in the performance evaluation: Please provide more details about data partitioning and the effect of skew. Also provide details about how star partitioning and a priori pruning contribute to performance. Please provide references to these two methods.}

\response { The evaluations are added in Table 7 and Figure 8. Due to the space limit, we only provide the summary of performance skew in Table 9 (i.e., the longest time-taking task vs stand deviation of each task). As per our analysis, the skew is limited because every stars are of similar sizes. The references of apriori and star partition are added accordingly in Section 5.1 and Section 5.2.
}