### 2016-08-26 ReviewComments/review1.tex

\section{Response to Reviewer 1}

\emph{The main con that I've spotted is the fact that the algorithm might have been
more nicely framed also within the Spark environment by taking advantage of
the various possibilities offered by it (e.g. caching of RDDs)}

\response{We agree that utilizing advanced Spark features would 
improve our solutions. We do not include these optimization in-depth
because in this work we set out focus on solving the challenges
in parallelizing the GCMP mining algorithms.
Nevertheless, we would improve and open-source a Spark-tailored
version of SPARE in the future.
}