### 2016-08-28 ReviewComments/review1.tex

\section{Response to Reviewer 1}

\emph{The main con that I've spotted is the fact that the algorithm might have been
more nicely framed also within the Spark environment by taking advantage of
the various possibilities offered by it (e.g. caching of RDDs)}

\response{We agree that utilizing advanced Spark features would 
improve our solutions. In this work, we do not include 
these optimization in-depth because we set out focus 
on solving the challenges in parallelizing the 
GCMP mining algorithms.
We are also in progress of open-sourcing a Spark-tailored
version of SPARE.
}