### 2016-08-28 ReviewComments/response.tex
New revision
...

\documentclass{vldb}
\usepackage{url}
\newcounter{ctResp}
\newcommand{\response}[1]{
\addtocounter{ctResp}{1}
\textbf{Response \arabic{ctResp}:} #1 \\}

\begin{document}
\title{Response to reviewers' comments for Paper \# 278}
\maketitle

We are deeply grateful to all the reviewers for the generous and insightful
suggestions which help us to improve this work to a better level.
In the following, we provide explicit response to each raised concern and address our corresponding revisions in the new submission.

\input{review1}
\input{review2}
\input{review3}
\input{meta}

\end{document}