### 2016-09-11 sec7_conclusion.tex

\vspace{-0.5em}
\section{Conclusions and Future Work}
\label{sec:concl}
In this paper, we proposed a generalized co-movement pattern to unify those proposed in the past literature. We also devised two types of parallel frameworks in Spark that can scale to support pattern detection in trajectory databases with hundreds of millions of points. The efficiency and scalability were verified by extensive experiments on three real datasets. In the future, we intend to examine co-movement pattern detection in streaming data for real-time monitoring. We are also interested in extending the current parallel frameworks to support other types of advanced patterns.