### 2016-10-28 Camera-Ready/sec2_relatedworks.tex

\section{Related Work}
\label{sec:related_works}
Related work can be grouped into three categories: \emph{co-movement patterns},
\emph{dynamic movement patterns} and \emph{trajectory mining frameworks}.
%The \emph{co-movement patterns} in literature consist 
%of five members, namely \emph{group}~\cite{wang2006grouppattern}, \emph{flock}~\cite{gudmundsson2004flock},
%\emph{convoy}~\cite{jeung2008convoy}, \emph{swarm}~\cite{li2010swarm} and \emph{platoon}~\cite{li2015platoon}.
%We have demonstrated the semantics of these patterns in Table~\ref{tbl:existing_co_patterns} and Figure~\ref{fig:related_work}. 
%In this section, we distinguish the parallel GCMP mining from these concepts.
%In this section, we focus on comparing the techniques used in these works.
%For more trajectory patterns other than \emph{co-movement patterns}, 
%interested readers may move to~\cite{zheng2015survey} for a comprehensive survey.
\subsection{Co-Movement Patterns}
\subsubsection{Flock and Convoy}
The difference between \emph{flock} and \emph{convoy} lies 
in the object clustering methods. In \emph{flock},
objects are clustered based on their distances. Specifically, the
objects in the same cluster need to have a pairwise distance less than \emph{min\_dist}. 
This essentially requires the objects to be within a disk-region of delimiter less than \emph{min\_dist}.
In contrast, \emph{convoy} clusters objects using density-based spatial clustering~\cite{ester1996density}.
Technically, \emph{flock} utilizes a $m^{th}$-order Voronoi diagram~\cite{laube2005finding} to detect whether
a subset of $n$ ($n \geq m$) objects stay in a disk region.
%a subset of object with size greater than $m$ stays in a disk-region.
\emph{Convoy} employs
a trajectory simplification~\cite{douglas1973linesimplification} technique to boost pairwise distance computations in
the density-based clustering.
After clustering, both \emph{flock} and \emph{convoy} use a sequential scanning
method to examine each snapshot. 
During the scan, the object groups that do 
not appear in consecutive snapshots are pruned.
%During the scan, object
%groups that appear in consecutive timestamps are detected. Meanwhile, the object groups that do not
%match the consecutive constraint are pruned. 
However, such a method faces high complexity when supporting other patterns.
For instance, in \emph{swarm}, the candidate set during the sequential scanning grows
exponentially, and many candidates can only be pruned after the entire dataset are scanned.

\subsubsection{Group, Swarm and Platoon}
Different from \emph{flock} and \emph{convoy}, all the \emph{group},\emph{swarm} and \emph{platoon}
patterns have more relaxed constraints on the pattern duration. Therefore, their techniques of mining are of
the same skeleton. The main idea of mining is to grow an object set from an empty set
in a depth-first manner. During the growth, various pruning techniques are provided to prune 
unnecessary branches. \emph{Group} pattern uses a VG-graph to guide the pruning of false candidates~\cite{wang2006grouppattern}.
\emph{Swarm} designs two more pruning rules called backward pruning and forward pruning~\cite{li2010swarm}. \emph{Platoon}~\cite{li2015platoon}
leverages a prefix table structure to steer the depth-first search, which shows efficiency 
as compared to the other two methods.
However, the pruning rules adopted by the three patterns heavily rely on depth-first search which loses efficiency in a parallel scenario.

\subsection{Other Related Trajectory Patterns}
A closely related literature to co-movement patterns is the \emph{dynamic movement} patterns. Instead of requiring the same set of object traveling together, \emph{dynamic movement} patterns allow objects to temporally join or leave a group. Typical works include \emph{moving clusters}~\cite{kalnis2005movingclusters}, \emph{evolving convoy}~\cite{aung2010discovery}, \emph{gathering}~\cite{zheng2013gathering} etc. These works cannot model GCMP since they enforce global consecutiveness on the 
temporal domain.

\subsection{Trajectory Mining Frameworks}
Jinno et al. in~\cite{jinno2012paralleltpattern} designed a MapReduce based algorithm to efficiently support \emph{T}-pattern discovery, where a \emph{T}-pattern is a set of objects visiting the same place at simliar time. Li et al. proposed a framework of processing online \emph{evolving group} pattern~\cite{li2013onlinegroup}, which focuses on supporting efficient updates of arriving objects. 
As these works essentially differ from co-movement pattern, their techniques cannot be directly applied to discover GCMPs.